% The name of this file: cddman.tex
% written by by Komei Fukuda
% December 14, 1993
% Revised March 15, 1994
%
\documentstyle[11pt]{article}
\renewcommand{\baselinestretch}{1.0}
\setlength{\oddsidemargin}{8mm}
\setlength{\textwidth}{16cm}
\setlength{\topmargin}{10mm}
\setlength{\textheight}{23cm}
\setlength{\headsep}{0in}
\setlength{\headheight}{0pt}

\pagestyle{empty}

\begin{document}

.\vspace{20mm}

\begin{center}

% {\LARGE RO931123}


\vspace{20mm}

{\LARGE cdd User Manual}

\vspace{20mm}

{\Large Komei Fukuda\\
   EPFL, Lausanne, Switzerland\\
   fukuda@dma.epfl.ch\\
   and\\
   University of Tsukuba, Tokyo, Japan\\
  fukuda@gssm.otsuka.tsukuba.ac.jp
}

\vspace{20mm}
{\Large (cdd Version 0.51c,  March 15, 1994)}

\end{center}

\newpage
\pagestyle{plain}
\setcounter{page}{1} 

\begin{center}

{\Large cdd User Manual}

\bigskip
{\large Komei Fukuda\\
   fukuda@dma.epfl.ch  or  fukuda@gssm.otsuka.tsukuba.ac.jp\\
   EPFL, Lausanne, Switzerland, and\\
   University of Tsukuba, Tokyo, Japan}

\bigskip
{\large (cdd Version 0.51c,  March 15, 1994)}

\end{center}

\section{Introduction} \label{INTRODUCTION}

The program  cdd.c  is a C implementation of 
the Double Description Method~\cite{MRTT}
for generating all vertices (i.e. extremal points)
and extremal rays of a general 
convex polyhedron given by a system of linear inequalities:
\[
   P = \{ x  \in R^d:  A  x  \le  b \}
\]
where $A$ is an $m \times d$ real matrix and $b$ is a real
$m$ dimensional vector.  It is intended that the program works 
only when the input polyhedron has at least one vertex but 
works for polyhedra $P$ which are not full-dimensional.    
 
This program has been produced
from the pascal code pdd.p (Version 0.2)
which was written by the author.   For the translation,
we applied the Pascal-to-C translator p2c
by Dave Gillespie to cdd.c, and we then modified
the C-code so that it runs without any p2c libraries.

One major advantage of this C-version over the pascal version is
that there is no restriction on the size  $m$ (and $d$), while in pdd.p
the size $m$ must be smaller than equal to the maximum element size
of SET TYPE variables in Pascal compiler used.

The program cdd reads input and writes output in 
{\em Polyhedra format\/} which was defined by David Avis and
the author.  The program called rs developed by David Avis is
a C-implementation of the reverse search algorithm~\cite{AF} 
for the same enumeration purpose, and it conforms to Polyhedra format as well.
Hopefully, this compatibility of the two programs
enables users to use both programs for the same input files
and to choose whichever is useful for their purposes.
From our experiences with relatively large problems,
the two methods are both useful and perhaps complementary
to each other.  In general, the program cdd tends to be
efficient for highly degenerate inputs and the program rs
tends to be efficient for nondegenerate or slightly
degenerate problems.

One interesting feature of cdd is its capability
of handling the dual (reverse)  problem without any transformation
of data.  The dual problem is known to be the 
{\em (convex) hull problem\/} which
is to obtain a linear inequality representation
of a convex polyhedron given as the Minkowski sum of 
the convex hull of a finite set of points and the nonnegative
hull of a finite set of points in $R^d$: 
$P = conv(v_1,\ldots,v_n) +  nonneg(r_1,\ldots,r_s)$.
As we see in this manual, the computation can be done
in straightforward manner.  There is one assumption for the input
for hull computation: the polyhedron must be full-dimensional.

Among the hardest problems that could be
solved by cdd is a 21-dimensional hull problem given by 63 
extremal rays. This polyhedral cone, known as the {\em complete
cut cone on 7 points\/}, has exactly 38,780 facets and
some of facets contain many rays, see e.g. \cite{G}. 
It took a couple of days for an earliear version of cdd
to compute the rays exactly on a fast workstation.
(Note that the present version 0.51 is considerabley faster.)  
The input file (ccc7.ine) of this polyhedra is
included in the distribution.

The size of an input file hardly indicates the degree of 
hardness of its vertex/ray enumeration.  While this program
can handle a highly degenerate problem (prodmT5.ine) with 
711 inequalities in  19 dimension quite easily with
the computation time 3-4 minutes on fast workstation, 
a 8-dimensional problem (mit729-9.ine) with 729 inequalities
can be extremely hard.  It took us about a month (for
an older version of cdd) on
a reasonably fast Sun 4 workstation.  The latter problem arises
from the ground state analysis of a ternary alloy model, see \cite{CGAF}.
Both input files are included in the distribution.  

Although the program can be used for nondegenerate inputs,
it might not be very efficient.  For nondegenerate inputs, 
the use of the reverse search code rs is strongly
recommended.

One serious drawback of the current cdd is that the computation
is done with floating point arithmetic, and it does
perform infinite precision arithmetic. In contrast, the program rs
supports multiple precision arithmetic with any fixed
multiplicity.  We are now developing
a new version of cdd, which will hopefully perform the exact
computation for integer and rational inputs.

This program can be distributed freely under the GNU GENERAL PUBLIC LICENSE.
Please read the file COPYING carefully before using.

I will not take any responsibility of any problems you might have
with this program.  But I will be glad to receive bug reports or suggestions
at the e-mail addresses above. 

\section{Polyhedra Input and Output Format} \label{FORMAT}
\bigskip
Polyhedra input format is quite simple:

\begin{tabular}{ccl}
\\ \hline
\multicolumn{3}{l} {various comments}\\
\multicolumn{3}{l} {\bf begin}\\
 $m$ & $d+1$ & {\bf numbertype}\\
 $b$ & $-A$ \\
\multicolumn{3}{l} {\bf end}\\
\multicolumn{3}{l} {various options} \\ \hline
\end{tabular}

\bigskip
\noindent
where numbertype can be one of integer, rational or real.
There is one restriction: the last $d$ rows must determine
a vertex of $P$.  The program cdd does not care whether
this condition is satisfied, as long as the polyhedron
contains at least one vertex.  But one can specify that 
the last $(d+1)$ rows be chosen as the initial set of
$(d+1)$ rows for the double description algorithm.
See {\bf initbasis\_at\_bottom} option in Section~\ref{OPTIONS}.

Polyhedra output format is quite similar to input format:

\begin{tabular}{ccl}
\\ \hline
\multicolumn{3}{l} {various comments produced by a program}\\
\multicolumn{3}{l} {\bf begin}\\
 $n+s$ & $d+1$ & {\bf numbertype}\\
 $1$ & $v_1$  & \\
 $\vdots$ & $\vdots$  & \\
 $1$ & $v_n$  & \\
 $0$ & $r_1$  & \\
 $\vdots$ & $\vdots$  & \\
 $0$ & $r_s$  & \\
\multicolumn{3}{l} {\bf end}\\  \hline
\end{tabular}

\bigskip
\noindent
where $v_1, \ldots, v_n$ are the vertices and
$r_1, \ldots, r_s$ are the extremal rays of the
polyhedron $P$.  Here we do not require that
the vertex list and the ray list are output
separately; they can appear mixed in arbitrary
order.


For example, let $P$ be the following unbounded 3-dimensional 
polyhedron given by
\[
   P = \{ x  \in R^3:
    1\le x_1 \le 2, \; 1 \le x_2 \le 2, \; 1 \le x_3\},
\]
which is a 3-cube without one ^^ ^^ lid". For finding all
vertices and extremal rays, the input file for cdd is

\begin{verbatim}
file name: ucube.ine
3 cube without one "lid"
begin
        5      4    integer
  2   -1   0   0
  2    0  -1   0
 -1    1   0   0
 -1    0   1   0
 -1    0   0   1
end
incidence
adjacency
\end{verbatim}

The meaning of options ^^ ^^ incidence" and ^^ ^^ adjacency" 
will be explained in Section~\ref{OPTIONS}.
After you run cdd with this input file, you will get
an output file, say ucube.ext, which looks like:

\begin{verbatim}
* cdd: Double Description Method C-Code:Version C0.51 (February 12, 1994)
* Copyright (C) 1994, Komei Fukuda, fukuda@dma.epfl.ch
*Input File:ucube.ine   (   5  x    4)
*HyperplaneOrder: LexMin
*Degeneracy preknowledge for computation: None (possible degeneracy)
*Vertex/Ray enumeration is chosen.
*Computation starts     at Mon Feb 14 16:13:27 1994
*            terminates at Mon Feb 14 16:13:27 1994
*Total processor time = 0 seconds
*                     = 0 hour 0 min 0 sec
*FINAL RESULT:
*Number of Vertices =       4,   Rays =       1
begin
        5      4    real
  1  2  1  1
  1  1  1  1
  1  1  2  1
  1  2  2  1
  0  0  0  1
end
\end{verbatim}

The output shows that the polyhedron has four vertices
$(2,1,1)$, $(1,1,1)$, $(1,2,1)$, $(2,2,1)$ and
only one extremal ray $(0,0,1)$.  The comments contain
information on the name of input file, and the options
chosen to run the program which will be explained in
the next section.

\newpage
\section{Options}  \label{OPTIONS}

The following options are available for cdd.  These options are
set if they appear in input file after the ^^ ^^ end" command.
Independent options can be set simultaneously, but each option
must be written separately in one line, and  two options
should not be written in one line.

\begin{description}

\item[hull] option \\
When this option is chosen, the program cdd
will do the reverse operation.   That is, the input is assumed to
be a set of points and directions (rays).  When this option is set,
it is required that each data line must start with 
either ^^ ^^ 1" or ^^ ^^ 0", meaning points and rays, respectively.
More specifically, if the input file is of form

\begin{tabular}{ccl}
\\ \hline
\multicolumn{3}{l} {comments}\\
\multicolumn{3}{l} {\bf begin}\\
 $n+s$ & $d+1$ & {\bf numbertype}\\
 $1$ & $v_1$  & \\
 $\vdots$ & $\vdots$  & \\
 $1$ & $v_n$  & \\
 $0$ & $r_1$  & \\
 $\vdots$ & $\vdots$  & \\
 $0$ & $r_s$  & \\
\multicolumn{3}{l} {\bf end}\\
\multicolumn{3}{l} {\bf hull} \\ \hline
\end{tabular}

\noindent
Then the input is interpreted as the polyhedron in $R^d$:\\
$P = conv(v_1,\ldots,v_n) +  nonneg(r_1,\ldots,r_s)$ 
and the output will be a minimal system of linear inequalities
to represent $P$. 

\item[dynout\_off] option\\
When this option is chosen, the program will
not output vertices and rays to the CRT in real time.  
The default is dynout\_on.

\item[stdout\_off] option\\
 When this option is chosen, the program will not
output any progress report of computation (iteration number. etc).
The default is stdout\_on.

\item[logfile\_on] option\\
 When this option is chosen, the program will output
to a specified file (*.ddl) some information on the computation history.
This can be useful when the user does not know which hyperplane order
(mincutoff, maxcutoff, mixcutoff, lexmin, lexmax, minindex)
is efficient for computation.

\item[incidence, \#incidence] options\\
When the {\bf incidence} option is selected, the incidence relations between
the vertices/rays and the inequalities will be output
in a separate file (*.icd).  Here, a vertex is said to be 
{\em incident with\/} an inequality if the inequality is satisfied by equality.
An extremal ray $r$ is said to be {\em incident with\/} 
an inequality $a^T \; x \le b$ if  $a^T \; r = 0$.  For example,
since the incidence option was set for the example input file ucube.ine in
the previous section, the program outputs the following ucube.icd file:
\begin{verbatim}
*Incidences of output(=vertices/rays) and input (=hyperplanes)
*   for each output, #incidence and the set of hyperplanes containing it
*   or its complement with its cardinality with minus sign
*cdd input file : ucube.ine   (   5  x    4)
*cdd output file: ucube.ext
begin
       5    5    6
       3    1    4    5
       3    3    4    5
       3    2    3    5
       3    1    2    5
      -1    5
end
\end{verbatim}
After ^^ ^^ begin", there are three numbers $5 \quad 5 \quad 6$.
The first number $5$ is a number of output (vertices and rays).
The next number $5$ is $m$, the number of inequalities in the input file.
The last number $6$ is usually $m+1$, and $m$ if the input linear inequality
system is homogeneous (i.e., has zero RHS) or the hull option is chosen.
The number $m+1$ corresponds to the infinity constraint which is added
for vertex/ray enumeration when the input system is not homogeneous.

The incidence data starts right after these three numbers.
At each line, the cardinality of incident inequalities and
the list of their indices are given.  There is an exception that, when
there are more incident inequalities than non-incident ones, then the program
outputs the list of non-incident inequalities with its
size with negative sign.  This is to save space of output.

For example, the first output line $3 \quad 1 \quad 4 \quad 5$ 
corresponds to the
first vertex of ucube.ext file in previous section, that is, 
the vertex $(2, 1, 1)$.  The first number $3$ is simply the number
of incident inequalities and the rest is the indices of
those inequalities, and so the 1st, 4th and 5th inequalities are
satisfied by equality at this vertex.   The last output
$-1 \quad 5$ corresponds to the ray  $(0,0,1)$.  Since all inequalities
except the last (5th) inequality are incident with this ray,
the output is the complementary list with its cardinality (=1) with negative
sign.  Note that the original list is $5 \quad 1 \: 2 \: 3 \: 4 \: 6$, where
$6$ is the infinity plane.  One can ignore the infinity
plane for some purposes,  but for analyzing the combinatorial 
structure of polyhedra, it is
very important information.

The {\bf \#incidence} option can be used when you do not wish
to output the incidence file but to output only the cardinality of incidence
for each output, at the end of each output line.


\item[nondegenerate] option\\
When this option is set, the program assumes that the input system
is not degenerate, i.e., there is no point in the space $R^d$ satisfying
more than $d$ inequalities of input with equality.
It will run faster with this option, but of course, 
if this option is set for degenerate inputs, it is 
quite possible that the output is incorrect.  
The default is this option being off.

\item[adjacency] option\\
This option can be used when you want to output the adjacency of output.
When the output is the list of vertices and rays, the program will
output the adjacency list.  For the example input ^^ ^^ ucube.ine",
the following extra file, say ^^ ^^ ucube.adj", will be created:
\begin{verbatim}
*Adjacency List of output (=vertices/rays)
*cdd input file : ucube.ine   (   5  x    4)
*cdd output file: ucube.ext
begin
  5
 1  3 : 2 4 5
 2  3 : 1 3 5
 3  3 : 2 4 5
 4  3 : 1 3 5
 5  4 : 1 2 3 4
end
\end{verbatim}
The first number $5$ is simply the number of outputs of cdd, the
number of vertices and rays in this case.
The second line $ 1 \quad  3 \quad : 2 \quad  4 \quad  5$ says 
that the first output of
ucube.ext file has degree (valency) $3$, and its three neighbours are
2nd, 4th and 5th output.

When the computation is to obtain the hull (inequality system),
the adjacency is of course that of inequalities (i.e. facets).

\item[mincutoff, maxcutoff, mixcutoff, lexmin, lexmax, minindex] options\\
The double description is an incremental algorithm which
computes the vertices/rays of a polyhedron given by some $k$ of
original inequalities from the precomputed vertices/rays of a
polyhedron given by $k-1$ inequalities.  It is observed that
the efficiency of the algorithm depends strongly on how
one selects the ordering of inequalities, although a little
can be said theoretically.
These options are to select the ordering of inequalities to be
added at each iteration, and it is recommended to do small
experiment to select good ordering for a specific type of problems.
Unfortunately, a good ordering depends on the problem and there does not seem 
to be THE BEST ordering for every computation.  From our experiences,
lexmin, lexmax, mincutoff, maxcuoff work quite well in general.


The default is lexmin ordering which simply order inequalities
with respect to lexico-graphic ordering of rows of $A$.  The lexmax
is reverse of lexmin.  The mincutoff (maxcutoff) option selects an inequality which
cuts off the minimum (maximum) number of vertices/rays of the $k-1$th polyhedron. 
The mixcutoff option is the mixture of mincutoff and maxcutoff which selects
an inequality which cuts off the $k-1$th polytope as unbalanced as possible.
The maxcutoff option might be efficient if the input contains
many redundant inequalities (many interior points for hull computation).
Finally, the minindex option selects the hyperplanes from the top of
the input.

\item[initbasis\_at\_bottom] option\\
When this option is set, the program tries to select
the initial set of rows for the double description
method from the bottom of the input.  This means that
if the last (d+1) rows are independent, 
they will be chosen to initiate the algorithm.

This option is {\em not\/} default. The default
follows the same ordering as the ordering
of inequalities chosen.  This means that if {\bf lexmin\/}
is the ordering of inequalities, then the initial 
independent rows
will be chosen sequentially with lexico-min ordering.
There are exceptions when this rule
is not applicable, i.e. when one of mincutoff or maxcutoff
options is chosen. In such cases, {\bf lexmin\/}
ordering will be chosen.

\item[maximize, minimize] option\\
When maximize option is set with an objective vector 
$c_0\: c_1 \: c_2 \ldots c_d$, the program
simply solves the linear program: $\max c_0 + c_1 x_1 + c_2 x_2 +\cdots + c_d x_d$
over the input polyhedron $P$. The grammer is simply

\begin{tabular}{ccl}
\\ \hline
\multicolumn{3}{l} {various comments}\\
\multicolumn{3}{l} {\bf begin}\\
 $m$ & $d+1$ & {\bf numbertype}\\
 $b$ & $-A$ \\
\multicolumn{3}{l} {\bf end}\\
\multicolumn{3}{l} {maximize} \\ 
\multicolumn{3}{l} { $c_0 \quad c_1 \quad c_2 \quad \cdots \quad c_d$ } \\ \hline
\\
\end{tabular}

The minimize option works exactly same way for minimization of
a linear objective function.
See the sample input file ^^ ^^ lptest.ine".  The program cdd
will output both primal and dual optimal solutions  if the LP
is solvable.  If the LP is infeasible (dual infeasible), then
it will output an evidence.  Currently these options are
not supported for homogeneous input (i.e. zero righthand side case).

\item[find\_interior] option\\
When this option is set, the program solves the linear program: 
$\max x_{d+1}$
subject to $ A x + e x_{d+1} \le b$, where $e$ is the
column vector of all $1$'s.  If the optimum value is zero,
the polyhedron has no interior point.  If the optimum value is 
negative then the polyhedron is empty. If the LP is dual inconsistent,
then the polyhedron admits unbounded inscribed balls.  To find any
interior point in this last case, one must add some inequality(ies)
to bound the polyhedron. 


\item[partial\_enumeration] option\\
With this option, one can enumerate only those
vertices and rays that are lying on the set of hyperplanes
associated with specified inequality numbers. If you want
to compute all vertices/rays lying on hyperplanes
associated with $k$ inequalities $i_1, i_2, \ldots, i_k$
($1 \le i_j \le m$), then
the option should be specified as

\begin{tabular}{ccl}
\\ \hline
\multicolumn{3}{l} {various comments}\\
\multicolumn{3}{l} {\bf begin}\\
 $m$ & $d+1$ & {\bf numbertype}\\
 $b$ & $-A$ \\
\multicolumn{3}{l} {\bf end}\\
\multicolumn{3}{l} {partial\_enumeration} \\ 
\multicolumn{3}{l} { $k \quad i_1 \quad i_2 \quad \cdots \quad i_k$ } \\ \hline
\\
\end{tabular}


See the input file ^^ ^^ partialtest.ine".

\item[preprojection] option\\
This option is for a preprocessing
 of orthogonal projection of the polyhedon to a subset of variables.  
That is, if the inequality inequality system is of two-block form  $A1 x1 + A2 x2 \le b$, 
and the variable indices for $x2$, say $1, 4, 6, 7$,
are listed in the input file as
\begin{verbatim}
begin
  m   d+1   numbertype
  b   -A1  -A2
end
preprojection
  4  1  4  6  7
\end{verbatim}
    
Then, cdd will output the inequality system,  $A1 x1  \le b$, together with the list $R$ of extremal
rays of the homogeneous cone  $\{z:  z  \ge 0  \mbox{ and }  z A2 = 0 \}$.  
Consequently, the inequality system
$\{ \quad r A1  x1  \le  r b : \quad r \in  R \}$
represents the projection of the original polyhedron onto
x1-space with possible redundancy.   There is a supplementary  C program, 
written by F. Margot, EPFL, which generates quickly a minimal
(i.e. nonredundant) system from these two outputs.
We are now preparing to distribute this program with cdd.
\end{description}

\section{How to Use}  \label{HOWTO}

The program hardly has any user interface.  Once you have a compiled
executable file, say {\em cdd\/} (see Section~\ref{CAUTIONS}), 
and once you create an input file,
say, {\em test.ine\/}, you have basically two ways to run the program.
The simplest way is just to run the program with 
\begin{verbatim}
  % cdd test.ine
\end{verbatim}
and then the program will open necessary output files with
default file names, and output the requested results.
The default names are {\em test.ext\/}, {\em test.icd\/}, {\em test.adj\/},
{\em testsub.ine\/} for the extremal points/ray file, the incidence
file, the adjacency file and the preprojection variable sub-inequality
system, respectively.  If you want to specify the output file names
different from default, simply run the program by
\begin{verbatim}
  % cdd
\end{verbatim}
and input desired file names at each of file name requests.

\section{Source Files and Compilation}  \label{CAUTIONS}

\begin{itemize}
\item[(1)] [Files and Compilation] The source files for distribution are

\begin{tabular}{ll}
     cdd.readme  &   This file itself\\
     cdd.c       &   C main source file\\
     cddarith.c  &   C sub source file\\  
     cdd.h       &   The header file for cdd.c\\
     cdddef.h    &   cdd definition file (whose two lines are to be edited by user)\\
     setoper.c   &   C library for set operation\\
     setoper.h   &   The header file for setoper.c \\
     cddman.tex  &  Latex source file of cdd User Manual\\
     cddHISTORY  &   brief description of changes made at each updates\\
     ine         &   A subdirectory containing sample input files\\
     ext         &   A subdirectory containing sample output files\\
     COPYING     &   GNU GENERAL PUBLIC LICENSE
\end{tabular}

\noindent
To compile the code in standard unix environment, type
\begin{verbatim}
  cc -O cdd.c cddarith.c setoper.c
\end{verbatim}

to obtain an optimized executable file a.out. The GNU compiler
gcc can replace cc above.
Since the program includes some standard ANSI library 
headers such as stdlib.h and time.h at compilation, 
the compiler must know the locations of the standard ANSI libraries.
Also, the files cdd.c, cdd.h, cdddef.h, cddarith.c, setoper.c and setoper.h are supposed to be
in the current directory.

\item[(2)] [Recompilation] The first two constants in the program cdddef.h are to be 
changed by the user if necessary, and the program must be recompiled
each time after any change is made.   These constants are simply
to specify the largest size of acceptable input data $(b, -A)$: 
\begin{verbatim}
#define MMAX    4001  /* USER'S CHOICE: max row size of A plus one */
#define NMAX    101   /* USER'S CHOICE: max column size of A plus one */
\end{verbatim}
If this input data has $m$ rows and $d+1$ columns, then in the program,
MMAX should be at least $m+1$ and NMAX should be at least
$d+1$.  Although it has no sense to set the sizes MMAX and NMAX much larger
than necessary, the program only creates spaces for MMAX+NMAX pointers
and uses only necessary storage space for each input, and
thus large MMAX and NMAX won't be too harmful.

Unlike the pascal version pdd, one can set the size MMAX as large as one
wants.  It is no more restricted by the SET TYPE element sizes of 
usual Pascal compilers.

\item[(3)] [TURBO/THINK C Users] The program cdd.c is written in
ANSI C, and thus it should run on
personal computers without any changes if one uses a compliler supporting
ANSI standard. We have verified that it can be compiled by THINK C without
any problems if ANSI library is added to the project together with
cdd.c, cddarith.c and setoper.c.  We have not yet tried to compile it with TURBO C.

\item[(4)] [Rational Input Not Supported] The present cdd.c does not
read any RATIONAL numbers ($a/b$ type with $a$, $b$ integers).

\end{itemize}

\section{FTP site}  \label{FTP}
An anonymous ftp site for the programs is set at EPFL:
\begin{verbatim}
   ftpsite:  ftp.epfl.ch (128.178.139.3)
   directory: incoming/dma
   filename: cdd-***.tar.Z
\end{verbatim}
Since the file is compressed binary file, it is necessary to use binary mode for
file transfer.

\bigskip
\noindent
{\large {\bf Acknowledgements.}} 
I am  grateful for Prof. Th. M. Liebling of
Department of Mathematics, EPFL, Switzerland who
provided me with an ideal opportunity to visit the department
for the academic year 1993-1994.  Without his 
generous help and encouragement,
the present form of this software would not have existed.
Also I am indebted to Dr. Alain Prodon and Mr. Francois Margot
of EPFL who used earlier versions of the program extensively,
found several bugs, and  gave me very helpful suggestions.
I am thankful to Dr. Henry Crapo of INRIA, France for
his precious advise on the input and output style of Polyhedra format.

\addtolength{\baselineskip}{-0.5\baselineskip}
\begin{thebibliography}{99}
\bibitem[AF92]{AF} Avis, D.  and Fukuda, K., ^^ ^^ A pivoting algorithm
        for convex hulls and vertex enumeration of arrangements
        and polyhedra," {\em Discrete Comput Geometry} 8
        (1992), pp. 295-313.

\bibitem[CGAF93]{CGAF} Ceder, G., Garbulski, G.D.,  Avis, D. and Fukuda, K., 
^^ ^^ Ground states of a ternary lattice model with nearest
and next-nearest neighbor interactions,"
{\em Physical Review B\/}, Vol.49, No. 1, (1994)  pp. 1-7.

\bibitem[G90]{G} Grishukhin, V.P., ^^ ^^ All facets of
the cut cone for n=7 are known," European Journal of
Combinatorics 11 (1990), 115-117.

\bibitem[MRTT53]{MRTT} Motzkin, T.S. , Raiffa, H.,  
Thompson, G.L. and  Thrall, R.M., ``The double description method," 
in  ^^ ^^ Contribution to the Theory of Games, Vol. II"
(H.W. Kuhn and A.W. Tucker, eds.),
Annals of Math. Studies 28,  Princeton University Press, 1953, pp.81-103.

\end{thebibliography}


\end{document}
